### 2018-02-18 doc/manuscript.Rmd
remove text related to pre-diabetes, num in promise, etc

- TGFA: Triacylglycerol fatty acids
- IS: Insulin sensitivity
- PROMISE: Prospective Metabolism and Islet Cell Evaluation cohort
- ISI: Insulin sensitivity index
- GEE: Generalized estimating equations
- PLS: Partial least squares
- HOMA2-%S: Homeostatic model of assessment 2 - percent sensitivity
- IGI/IR: Insulinogenic index over homeostatic model of assessment for insulin
resistance
- ISSI-2: Insulin Secretion-Sensitivity Index-2
- DNL: De novo lipogenesis
- ALT: Alanine aminotransferase
- NEFA: Non-esterified fatty acids
- OGTT: Oral glucose tolerance test
- CV: Coefficient of variation
- IFG: Impaired fasting glucose
- IGT: Impaired glucose tolerance
- MAQ: Modified activity questionnaire
- WC: Waist circumference
- MET: Metabolic equivalent of task
- SD: Standard deviation

@@

```{r inline}
dys <- project_data %>% 
    prep_dys_data() %>% 
    calc_convert_dys()

tagfa <- calc_tagfa_percent()
outcome_chg <- calc_outcome_changes()
female <- calc_discr_npct(ds0yr$Sex, 'Female')
euro <- calc_discr_npct(ds0yr$Ethnicity, 'European')
fhd <- calc_discr_npct(ds0yr$FamHistDiab, 'Yes')
gmest0 <- calc_gee_estci(gee_results$unadj)
gmest1 <- calc_gee_estci(gee_results$adj)
fa_cor <- calc_cor(cor_results$conc)
fa_cor_mol <- calc_cor(cor_results$pct)
cor_pls_homa2 <- calc_pred_corr(pls_homa2_t, pls_homa2_t$test_data)
cor_pls_isi <- calc_pred_corr(pls_isi_t, pls_isi_t$test_data)
cor_pls_igi <- calc_pred_corr(pls_igi_t, pls_igi_t$test_data)
cor_pls_issi2 <- calc_pred_corr(pls_issi2_t, pls_issi2_t$test_data)
```

@@

After 8-12 hours of overnight fasting, participants completed a 75g oral glucose tolerance test (OGTT) at 
each examination visit, with blood samples taken at fasting, 30 min, and 2 hr 
post-glucose load. Samples were subsequently processed and frozen at -70°C.
Alanine aminotransferase (ALT) was measured using standard laboratory 
procedures. Cholesterol, HDL, and clinically-measured TG 
were measured using Roche Modular's enzymatic colorimetric tests (Mississauga, 
ON). Both insulin and glucose were
measured from OGTT blood samples at fasting, 30 minute,
and 2 hour time points. Specific insulin was measured with the Elecsys 1010 
(Roche Diagnostics, Basel, Switzerland) immunoassay analyzer and 
electrochemiluminescence immunoassay, which shows 0.05% cross-reactivity to 
intact human pro-insulin and the Des 31,32 circulating split form (Linco Res. 
Inc) and has a coefficient of variation (CV) of 9.3%. Glucose was determined 
using an enzymatic hexokinase method (Roche Modular, Roche Diagnostics) with a 
detection range of 0.11 to 41.6 mmol/L and an inter-assay CV of <1.1% and an
intra-assay CV of < 1.9%.  All assays were performed at the Banting and Best
Diabetes Centre Core Lab at Mt Sinai Hospital. Impaired fasting glucose (IFG),
impaired glucose tolerance (IGT), and diabetes were categorized using the 2006
WHO criteria [@WHO2006].

@@

`r cite_t('basic')` shows basic characteristics of the PROMISE cohort. The mean 
follow-up time was `r calc_followup_time()` years, where `r calc_pct_full_visits()`%
of participants attended all three visits. There were
`r female$npct` females and `r euro$npct` who had European-ancestry, with a mean
age in years of `r aide::ave_sd(ds0yr$Age)` and a mean BMI
of `r aide::ave_sd(ds0yr$BMI)` kg/m^2^. As expected from the study's eligibility
criteria, the majority of participants, n=`r fhd$npct`, had a family history of
diabetes. Between the baseline visit and the 6-year visit in this sample,
insulin sensitivity and beta-cell function measures had a significant median decline of
between `r outcome_chg$chg` (p`r outcome_chg$p` from GEE; n=`r outcome_chg$n`). There 
were `r dys$pre` and `r dys$dm` incident cases of pre-diabetes (IFG and IGT) and
diabetes, respectively, over the 6-years; these observations were excluded from GEE and PLS
analyses due to how the outcomes were measured (OGTT was not done on those with diabetes).

>>>>

- TGFA: Triacylglycerol fatty acids
- IS: Insulin sensitivity
- PROMISE: Prospective Metabolism and Islet Cell Evaluation cohort
- ISI: Insulin sensitivity index
- GEE: Generalized estimating equations
- PLS: Partial least squares
- HOMA2-%S: Homeostatic model of assessment 2 - percent sensitivity
- IGI/IR: Insulinogenic index over homeostatic model of assessment for insulin
resistance
- ISSI-2: Insulin Secretion-Sensitivity Index-2
- DNL: De novo lipogenesis
- ALT: Alanine aminotransferase
- NEFA: Non-esterified fatty acids
- OGTT: Oral glucose tolerance test
- CV: Coefficient of variation
- MAQ: Modified activity questionnaire
- WC: Waist circumference
- MET: Metabolic equivalent of task
- SD: Standard deviation

@@

```{r inline}
tagfa <- calc_tagfa_percent()
outcome_chg <- calc_outcome_changes()
female <- calc_discr_npct(ds0yr$Sex, 'Female')
euro <- calc_discr_npct(ds0yr$Ethnicity, 'European')
fhd <- calc_discr_npct(ds0yr$FamHistDiab, 'Yes')
gmest0 <- calc_gee_estci(gee_results$unadj)
gmest1 <- calc_gee_estci(gee_results$adj)
fa_cor <- calc_cor(cor_results$conc)
fa_cor_mol <- calc_cor(cor_results$pct)
cor_pls_homa2 <- calc_pred_corr(pls_homa2_t, pls_homa2_t$test_data)
cor_pls_isi <- calc_pred_corr(pls_isi_t, pls_isi_t$test_data)
cor_pls_igi <- calc_pred_corr(pls_igi_t, pls_igi_t$test_data)
cor_pls_issi2 <- calc_pred_corr(pls_issi2_t, pls_issi2_t$test_data)
```

@@

After 8-12 hours of overnight fasting, participants completed a 75g oral glucose tolerance test (OGTT) at 
each examination visit, with blood samples taken at fasting, 30 min, and 2 hr 
post-glucose load. Samples were subsequently processed and frozen at -70°C.
Alanine aminotransferase (ALT) was measured using standard laboratory 
procedures. Cholesterol, HDL, and clinically-measured TG 
were measured using Roche Modular's enzymatic colorimetric tests (Mississauga, 
ON). Both insulin and glucose were
measured from OGTT blood samples at fasting, 30 minute,
and 2 hour time points. Specific insulin was measured with the Elecsys 1010 
(Roche Diagnostics, Basel, Switzerland) immunoassay analyzer and 
electrochemiluminescence immunoassay, which shows 0.05% cross-reactivity to 
intact human pro-insulin and the Des 31,32 circulating split form (Linco Res. 
Inc) and has a coefficient of variation (CV) of 9.3%. Glucose was determined 
using an enzymatic hexokinase method (Roche Modular, Roche Diagnostics) with a 
detection range of 0.11 to 41.6 mmol/L and an inter-assay CV of <1.1% and an
intra-assay CV of < 1.9%.  All assays were performed at the Banting and Best
Diabetes Centre Core Lab at Mt Sinai Hospital.

@@

`r cite_t('basic')` shows basic characteristics of the PROMISE cohort. The mean 
follow-up time was `r calc_followup_time()` years, where `r calc_pct_full_visits()`%
of participants attended all three visits. There were
`r female$npct` females and `r euro$npct` who had European-ancestry, with a mean
age in years of `r aide::ave_sd(ds0yr$Age)` and a mean BMI
of `r aide::ave_sd(ds0yr$BMI)` kg/m^2^. As expected from the study's eligibility
criteria, the majority of participants, n=`r fhd$npct`, had a family history of
diabetes. Between the baseline visit and the 6-year visit in this sample,
insulin sensitivity and beta-cell function measures had a significant median decline of
between `r outcome_chg$chg` (p`r outcome_chg$p` from GEE; n=`r outcome_chg$n`).
